% =========================================================
% Ordinary Differential Equations --- Breadcrumb
% =========================================================

\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where This Chapter Lives}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Calculus
$\;\to\;$ \textbf{Ordinary Differential Equations}
$\;\to\;$ Dynamical Systems / Numerical Methods / PDE
\end{center}

\medskip
\noindent\textbf{Why this chapter belongs here.}
Ordinary differential equations are where derivatives stop being only local
rates and start governing whole families of evolving states.  Once
differentiation is in hand, equations such as
\[
y'(t)=f(t,y(t))
\]
become the natural language for motion, growth, decay, forcing, and
feedback.

\medskip
\noindent\textbf{What this chapter will gather.}
The eventual ODE chapter is the right home for:
\begin{itemize}
  \item first-order equations and separable models;
  \item linear equations and integrating factors;
  \item existence and uniqueness ideas;
  \item systems of ODEs and phase portraits;
  \item stability, equilibrium, and qualitative behavior;
  \item numerical methods such as Euler and Runge--Kutta schemes.
\end{itemize}

\medskip
\noindent\textbf{Why this is currently a breadcrumb.}
The surrounding curriculum is still building the supporting architecture:
computational calculus, linear algebra, and numerical analysis.  ODEs sit at
the meeting point of all three, so this file marks the destination clearly
even before the chapter is fully written.
\end{tcolorbox}

