% =========================================================
% Geometric Modeling — Parametric Curves and Surfaces
% =========================================================

\section{Parametric Curves and Surfaces}

% ---------------------------------------------------------
% Toolkit
% ---------------------------------------------------------

% ---------------------------------------------------------
\subsection*{Parametric Curves}

\begin{propositionbox}{Definition (Parametric Curve)}
A \emph{parametric curve} in $\mathbb{R}^n$ is a vector-valued function
\[
  C\,:\,[a,b]\;\to\;\mathbb{R}^n, \qquad
  C(u) = \bigl(x_1(u),\,x_2(u),\,\ldots,\,x_n(u)\bigr),
\]
where each coordinate function $x_i(u)$ is a real-valued function
of the scalar parameter $u$.
The interval $[a,b]$ is typically normalised to $[0,1]$.
\end{propositionbox}

\begin{propositionbox}{Definition (Velocity, Regularity, Unit Tangent)}
The \emph{velocity vector} (or \emph{tangent vector}) at $u$ is the derivative
\[
  C'(u) = \bigl(x_1'(u),\,x_2'(u),\,\ldots,\,x_n'(u)\bigr).
\]
The curve is \emph{regular} at $u$ if $C'(u) \neq \mathbf{0}$.
At a regular point, the \emph{unit tangent vector} is
\[
  \hat{T}(u) = \frac{C'(u)}{\|C'(u)\|}.
\]
The scalar $\|C'(u)\|$ is the \emph{speed} of traversal.
\end{propositionbox}

\begin{remark*}[Properties of parametric curves]
Parametric curves carry structure that implicit curves do not:
\begin{itemize}
  \item \textbf{Direction.}
        The parameter increases from $a$ to $b$, giving the curve
        a natural orientation.
        Reversing the parametrisation reverses the direction.
  \item \textbf{Velocity and speed.}
        $C'(u)$ measures both the direction of motion and how fast
        the curve is being traversed.
        A curve can be \emph{reparametrised} (change of variable
        $u = \phi(t)$) without changing its shape;
        the velocity changes but the trace does not.
  \item \textbf{Endpoint interpolation.}
        The start and end points are $C(a)$ and $C(b)$, accessible directly.
  \item \textbf{Boundedness.}
        The curve is defined on a closed interval, so it represents
        a bounded segment, not an infinite line.
  \item \textbf{Cusps.}
        Where $C'(u) = \mathbf{0}$, the curve may have a cusp
        (velocity drops to zero, direction may reverse).
        These are degenerate points excluded by the regularity condition.
\end{itemize}
\end{remark*}

% ---------------------------------------------------------
\subsection*{Parametric Surfaces}

\begin{propositionbox}{Definition (Parametric Surface)}
A \emph{parametric surface} in $\mathbb{R}^3$ is a vector-valued function
of two parameters,
\[
  S\,:\,R\;\longrightarrow\;\mathbb{R}^3, \qquad
  S(u,v) = \bigl(x(u,v),\;y(u,v),\;z(u,v)\bigr),
\]
where $R\subseteq\mathbb{R}^2$ is the \emph{parameter domain}
(typically the unit square $[0,1]^2$).
\end{propositionbox}

% ---------------------------------------------------------
\subsection*{Partial Derivatives and Isoparametric Curves}

\begin{propositionbox}{Definition (Partial Derivative Vectors)}
The \emph{partial derivative vectors} of $S$ at $(u,v)$ are
\[
  S_u(u,v) = \frac{\partial S}{\partial u}
  = \left(\frac{\partial x}{\partial u},\;
          \frac{\partial y}{\partial u},\;
          \frac{\partial z}{\partial u}\right),
  \qquad
  S_v(u,v) = \frac{\partial S}{\partial v}
  = \left(\frac{\partial x}{\partial v},\;
          \frac{\partial y}{\partial v},\;
          \frac{\partial z}{\partial v}\right).
\]
$S_u$ is the tangent vector to the surface in the $u$-direction:
it is the velocity of the curve obtained by fixing $v$ and varying $u$.
$S_v$ is the corresponding tangent vector in the $v$-direction.
\end{propositionbox}

\begin{propositionbox}{Definition (Isoparametric Curves)}
Fixing one parameter on $S(u,v)$ produces a curve on the surface:
\begin{itemize}
  \item \emph{$u$-curve} at $v = v_0$:\quad
        $C_{v_0}(u) = S(u,\,v_0)$,\quad tangent $= S_u(u,v_0)$.
  \item \emph{$v$-curve} at $u = u_0$:\quad
        $C_{u_0}(v) = S(u_0,\,v)$,\quad tangent $= S_v(u_0,v)$.
\end{itemize}
These \emph{isoparametric curves} (isocurves) sweep the surface and
form its parameter grid.
The two families intersect at every surface point $S(u_0,v_0)$.
\end{propositionbox}

% ---------------------------------------------------------
\subsection*{Directional Derivative}

\begin{propositionbox}{Definition (Directional Derivative on a Surface)}
The \emph{directional derivative} of $S$ at $(u_0,v_0)$ in the
direction $\mathbf{d} = (d_u, d_v)$ (a unit vector in parameter space) is
\[
  D_{\mathbf{d}}\,S = d_u\,S_u(u_0,v_0) + d_v\,S_v(u_0,v_0).
\]
It measures the rate of change of position on the surface as the
parameter moves in direction $\mathbf{d}$ in the $uv$-plane.
Every tangent vector to the surface at $(u_0,v_0)$ is a directional
derivative for some choice of $\mathbf{d}$.
\end{propositionbox}

% ---------------------------------------------------------
\subsection*{Unit Normal and Tangent Plane}

\begin{propositionbox}{Definition (Regularity, Unit Normal, Tangent Plane)}
The surface is \emph{regular} at $(u_0,v_0)$ if
$S_u \times S_v \neq \mathbf{0}$ there.

At a regular point, the \emph{unit normal vector} is
\[
  N(u_0,v_0)
  = \frac{S_u(u_0,v_0)\;\times\; S_v(u_0,v_0)}
         {\|S_u(u_0,v_0)\;\times\; S_v(u_0,v_0)\|}\,.
\]

The \emph{tangent plane} at $S(u_0,v_0)$ is the plane through
$S(u_0,v_0)$ spanned by $S_u$ and $S_v$:
\[
  \Pi\;=\;\bigl\{\,S(u_0,v_0) + \alpha\,S_u + \beta\,S_v
               \;\bigm|\; \alpha,\beta\in\mathbb{R}\,\bigr\}.
\]
Its equation in terms of position $\mathbf{p}$ is
\[
  N\cdot\bigl(\mathbf{p} - S(u_0,v_0)\bigr) = 0.
\]
\end{propositionbox}

\begin{remark*}[Geometric meaning of $S_u \times S_v$]
The magnitude $\|S_u \times S_v\|$ equals the area of the
parallelogram spanned by $S_u$ and $S_v$, which is the
\emph{local area element} of the surface.
Where $\|S_u \times S_v\| = 0$ the surface is degenerate
(the two tangent directions are parallel or one vanishes),
and the normal is not defined.
\end{remark*}

% ---------------------------------------------------------
% Figure
% ---------------------------------------------------------
\vspace{1em}
\begin{center}
\input{volume-vi/geometric-modeling/notes/curves-and-surfaces/figure-parametric-surface}
\end{center}

\begin{center}
\small\itshape
Figure: A parametric surface patch $S(u,v)$ with its parameter domain (lower left).
At the point $S(u_0,v_0)$, the partial derivative vectors $S_u$ (red) and $S_v$ (green)
are tangent to the isoparametric curves; together they span the tangent plane (orange).
The unit normal $N = S_u \times S_v\,/\,\|S_u \times S_v\|$ (blue) is perpendicular
to the tangent plane.
\end{center}

% ---------------------------------------------------------
% Status
% ---------------------------------------------------------
\vspace{1em}
\begin{tcolorbox}[
  colback=gray!6, colframe=gray!40, arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Status: Active}},
  fonttitle=\small\bfseries
]
\noindent This section is actively expanding alongside work in Piegl \& Tiller,
\textit{The NURBS Book}.
Subsequent sections on Bézier curves, B-splines, NURBS, and continuity
conditions follow.
\end{tcolorbox}
