% =========================================================
% Chapter: Calculus
% =========================================================
\chapter{Calculus}

\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Real Numbers
$\;\to\;$ \textbf{Calculus (Computational Fluency)}
$\;\to\;$ Real Analysis
$\;\to\;$ Multivariable Analysis
$\;\to\;$ Differential Equations
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent
Volume VI treats calculus as a working toolkit. The real number system is
already available; here the emphasis is on computing limits, rates of change,
accumulation, series, and multivariable quantities accurately and fluently.
\end{tcolorbox}

\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Structural Role: Calculus as Computational Fluency}},
  fonttitle=\small\bfseries
]
This chapter is a computational field manual. It records definitions,
computational theorem statements, rule tables, examples, method notes, and
exercises. Rigorous proofs of the main calculus theorems belong in the Real
Analysis volume.

\medskip
\noindent
The guiding dependency pattern is:
\[
  \text{Volume VI computational theorem}
  \longrightarrow
  \text{Volume II rigorous theorem}
  \longrightarrow
  \text{proof file}.
\]
\end{tcolorbox}

\section*{Structural Roadmap}

\begin{enumerate}
  \item Limits: numerical behavior, algebraic methods, one-sided behavior,
        infinity, squeeze arguments, and asymptotes.
  \item Continuity: pointwise and interval continuity, discontinuities, and
        the Intermediate Value Theorem as a computational tool.
  \item Derivatives: difference quotients, rule algebra, standard derivative
        tables, chain-rule patterns, implicit and logarithmic differentiation,
        and applications.
  \item Integrals: antiderivatives, definite integrals, the Fundamental
        Theorem of Calculus as a computational bridge, techniques, improper
        integrals, and applications.
  \item Sequences and series: convergence, standard tests, power series, and
        Taylor/Maclaurin expansions.
  \item Multivariable calculus: vectors, curves, partial derivatives,
        gradients, multiple integrals, line and surface integrals, and the
        vector-calculus theorems.
\end{enumerate}

\clearpage
\input{volume-vi/calculus/limits/index}
\clearpage
\input{volume-vi/calculus/continuity/index}
\clearpage
\input{volume-vi/calculus/derivatives/index}
\clearpage
\input{volume-vi/calculus/integrals/index}
\clearpage
\input{volume-vi/calculus/sequences-and-series/index}
\clearpage
\input{volume-vi/calculus/multivariable-calculus/index}

\clearpage
\section{Capstone}
\textit{To be populated.}
