% =========================================================
% Computational Linear Algebra — Notes Index
% =========================================================

% ---------------------------------------------------------
% Breadcrumb
% ---------------------------------------------------------
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Algebraic Structures
$\;\to\;$ Linear Algebra (Axler)
$\;\to\;$ \textbf{Computational Linear Algebra}
$\;\to\;$ Geometric Modeling
$\;\to\;$ Differential Geometry
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{Relationship to Axler.}
\textit{Linear Algebra Done Right} is the theoretical spine for this
material --- vector spaces, linear maps, spectral theory, all developed
rigorously and proof-first.
This chapter is deliberately the other half: computational fluency
with the same machinery.
The two chapters are complementary and should not be conflated.
Axler proves why the tools work; this chapter practises using them.

\medskip
\noindent\textbf{What this chapter builds.}
The computational toolkit needed to work through geometric modeling,
numerical algorithms, and applied mathematics: vector arithmetic,
matrix operations, solving linear systems, orthogonality, determinants,
eigenvalues, the SVD, linear transformations, and selected applications.
The structure follows Strang, \textit{Introduction to Linear Algebra}
(5th ed.), and sections are populated just-in-time as downstream work
demands them.

\medskip
\noindent\textbf{Current priority.}
Chapters~1, 2 (§2.4--2.5), 3 (§3.4), 5 (§5.2), and 8 (§8.1--8.2)
provide the linear algebra required for Piegl \& Tiller,
\textit{The NURBS Book}, Chapters~1--3.
The remaining sections are stubbed and will be populated as needed.
\end{tcolorbox}
\vspace{1em}

% ---------------------------------------------------------
% Structural Roadmap
% ---------------------------------------------------------
\subsection*{Structural Roadmap}

\begin{center}
\textbf{Computation $\longrightarrow$ Geometry $\longrightarrow$ Application}
\end{center}

The progression follows Strang's chapter sequence:

\begin{enumerate}
  \item Vectors and linear combinations
  \item Solving linear equations
  \item Vector spaces and subspaces
  \item Orthogonality
  \item Determinants
  \item Eigenvalues and eigenvectors
  \item The Singular Value Decomposition
  \item Linear transformations
  \item Complex vectors and matrices
  \item Applications
  \item Numerical linear algebra
  \item Linear algebra in probability and statistics
\end{enumerate}

\begin{remark*}[Primary source]
Strang, \textit{Introduction to Linear Algebra}, 5th ed.\ (2016);
lectures freely available via MIT OpenCourseWare, 18.06.
\end{remark*}

% ---------------------------------------------------------
% Section inputs (uncomment as content is written)
% ---------------------------------------------------------

\input{volume-vi/computational-linear-algebra/notes/vectors/index}
\clearpage
\input{volume-vi/computational-linear-algebra/notes/solving-linear-equations/index}
\clearpage
\input{volume-vi/computational-linear-algebra/notes/vector-spaces/index}
\clearpage
\input{volume-vi/computational-linear-algebra/notes/orthogonality/index}
\clearpage
\input{volume-vi/computational-linear-algebra/notes/determinants/index}
\clearpage
\input{volume-vi/computational-linear-algebra/notes/eigenvalues/index}
\clearpage
\input{volume-vi/computational-linear-algebra/notes/svd/index}
\clearpage
\input{volume-vi/computational-linear-algebra/notes/linear-transformations/index}
\clearpage
\input{volume-vi/computational-linear-algebra/notes/complex-vectors-matrices/index}
\clearpage
\input{volume-vi/computational-linear-algebra/notes/applications/index}
\clearpage
\input{volume-vi/computational-linear-algebra/notes/numerical-linear-algebra/index}
\clearpage
\input{volume-vi/computational-linear-algebra/notes/probability-statistics/index}
\clearpage
