% =========================================================
% Chapter: Computational Geometry
% =========================================================
\chapter{Computational Geometry}

\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Linear Algebra
$\;\to\;$ Geometric Modeling
$\;\to\;$ \textbf{Computational Geometry}
$\;\to\;$ Optimization
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{How we got here.}
Linear algebra supplied coordinates, vectors, dot products, and linear
functionals. Geometric modeling showed how finite control data can determine
geometric objects. Computational geometry keeps the finite data visible and
asks which geometric structures can be defined and constructed from it.

\medskip
\noindent\textbf{What this chapter builds.}
The chapter begins with convex hulls: the smallest convex set containing a
finite point set, its extreme points, supporting lines, monotone hull structure,
and the greedy construction of the upper hull.

\medskip
\noindent\textbf{Where this leads.}
Convex hulls connect finite geometry with order, maximization, separation, and
algorithmic construction. Later computational-geometry topics use the same
pattern: define the geometric object exactly, identify its certificates, and
then turn those certificates into algorithms.
\end{tcolorbox}

\clearpage
\input{volume-vi/computational-geometry/notes/index}
\clearpage

\section*{Proofs}
\input{volume-vi/computational-geometry/proofs/index}
\clearpage
