\subsection*{Structure Signatures}
\label{subsec:structure-signatures}

\begin{definitionbox}{Definition (First-Order Signature)}
\begin{definition}[First-Order Signature]
\label{def:first-order-signature}
A \textbf{first-order signature} $\mathcal{L}$ consists of constant symbols,
function symbols with specified arities, and relation symbols with specified
arities.
\end{definition}
\end{definitionbox}

\begin{remark*}[Standard quantified statement]
\[
\mathcal{L}=(\mathrm{Const}_{\mathcal{L}},
\mathrm{Func}_{\mathcal{L}},
\mathrm{Rel}_{\mathcal{L}},\operatorname{arity})
\]
where each function and relation symbol has a specified arity.
\end{remark*}

\begin{remark*}[Definition predicate reading]
A first-order signature is the syntactic list of symbols that a structure must
interpret.
\end{remark*}

% Omitted Negated quantified statement: a signature is specified data, not a
% predicate whose negation is useful here.

\begin{remark*}[Interpretation]
The signature describes what kinds of constants, operations, and relations are
visible before any particular carrier set is chosen.
\end{remark*}

\begin{dependencies}
\begin{itemize}
  \item \hyperref[def:nullary-operation]{Nullary Operation}
  \item \hyperref[def:unary-operation]{Unary Operation}
  \item \hyperref[def:binary-operation]{Binary Operation}
  \item \hyperref[def:binary-relation]{Binary Relation}
\end{itemize}
\end{dependencies}

\begin{definitionbox}{Definition ($\mathcal{L}$-Structure)}
\begin{definition}[$\mathcal{L}$-Structure]
\label{def:algebraic-l-structure}
Given a signature $\mathcal{L}$, an \textbf{$\mathcal{L}$-structure}
$\mathcal{A}$ consists of a nonempty carrier set $A$ together with
interpretations of each constant, function, and relation symbol of
$\mathcal{L}$ on $A$.
\end{definition}
\end{definitionbox}

\begin{remark*}[Standard quantified statement]
\[
\mathcal{A}\models\mathcal{L}
\Longleftrightarrow
\mathcal{A}=(A,\ldots)
\text{ interprets each symbol of }\mathcal{L}\text{ on }A.
\]
\end{remark*}

\begin{remark*}[Definition predicate reading]
$\mathcal{A}$ is an $\mathcal{L}$-structure exactly when it gives meanings to
the symbols of $\mathcal{L}$ on a carrier set.
\end{remark*}

\begin{remark*}[Negated quantified statement]
\[
\neg\operatorname{Structure}_{\mathcal{L}}(\mathcal{A})
\]
when some symbol of $\mathcal{L}$ is not interpreted with the required arity on
the carrier set.
\end{remark*}

\begin{remark*}[Interpretation]
The signature is the type; the structure is the instance.  A group signature
names multiplication, identity, and inverse, while a particular group
interprets those symbols on a carrier set.
\end{remark*}

\begin{dependencies}
\begin{itemize}
  \item \hyperref[def:first-order-signature]{First-Order Signature}
  \item \hyperref[def:carrier-set]{Carrier Set}
\end{itemize}
\end{dependencies}
