% =========================================================
% Geometric Modeling — Notes Index
% =========================================================

% ---------------------------------------------------------
% Breadcrumb
% ---------------------------------------------------------
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Analytical Geometry
$\;\to\;$ Linear Algebra
$\;\to\;$ \textbf{Geometric Modeling}
$\;\to\;$ Differential Geometry
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{How we got here.}
Analytical geometry gave algebraic descriptions of geometric
objects in coordinates.
Linear algebra provided the language of transformations.
Geometric modeling applies both to represent and manipulate
\emph{complex} geometric shapes --- curves and surfaces that
cannot be described by a single equation --- using piecewise
polynomial constructions.

\medskip
\noindent\textbf{What this chapter builds.}
B\'ezier curves and their de~Casteljau algorithm; B-splines and
NURBS; parametric surface patches; subdivision surfaces;
affine and projective transformations in homogeneous
coordinates; the relationship between geometric modeling
constructions and the underlying algebraic and differential
geometry; and curvature analysis of modeled curves and surfaces.

\medskip
\noindent\textbf{Where this leads.}
Geometric modeling is the computational face of differential
geometry.
The smooth curves and surfaces studied abstractly in differential
geometry are approximated and manipulated concretely here.
For the main curriculum trajectory, this chapter provides
applied grounding and a source of concrete examples before the
more abstract machinery of manifolds and Riemannian geometry.
\end{tcolorbox}
\vspace{1em}

% ---------------------------------------------------------
% Structural Role
% ---------------------------------------------------------
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Structural Role: Geometric Modeling — Geometry Made Computable}},
  fonttitle=\small\bfseries
]
Geometric modeling is where the abstract geometry sequence
meets computation.

B\'ezier curves, B-splines, and NURBS are not merely engineering
tools: they are polynomial and rational maps from parameter
space into $\mathbb{R}^n$, and every property of interest
(smoothness, curvature, convex hull containment) is a theorem
in disguise.
The de~Casteljau algorithm is a recursive application of
affine interpolation; subdivision surfaces are limit constructions
whose analysis uses the spectral theory of subdivision matrices.

This chapter serves two roles in the project.
First, it provides a concrete, computable bridge between the
algebra of polynomials and the geometry of curves and surfaces.
Second, it is an optional applied detour that reinforces the
differential-geometric concepts (parametrisation, arc length,
curvature) in a setting where explicit computations are feasible.
\end{tcolorbox}
\vspace{1em}

% ---------------------------------------------------------
% Structural Roadmap
% ---------------------------------------------------------
\subsection*{Structural Roadmap}

\begin{center}
\textbf{Definitions $\longrightarrow$ Main Theorems
$\longrightarrow$ Consequences and Structural Insight}
\end{center}

The global progression is:

\begin{enumerate}
  \item Parametric curves: the modeling perspective
  \item B\'ezier curves: Bernstein polynomials and the de~Casteljau
        algorithm
  \item Properties of B\'ezier curves: convex hull, variation
        diminishing, affine invariance
  \item B-splines: knot vectors, basis functions, and local support
  \item NURBS: rational extensions and conic sections
  \item Parametric surface patches: tensor-product and triangular
  \item Continuity conditions: $C^k$ and $G^k$ joins
  \item Subdivision curves and surfaces
  \item Curvature analysis of parametric curves and surfaces
\end{enumerate}

\begin{remark*}[Primary sources]
Farin, \textit{Curves and Surfaces for CAGD};
Gallier, \textit{Geometric Methods and Applications};
Prautzsch, Boehm \& Paluszny, \textit{B\'ezier and B-Spline
Techniques}.
\end{remark*}


\input{volume-vi/geometric-modeling/notes/curves-and-surfaces/index}
\clearpage
% Future sections (uncomment as content is written):
\input{volume-vi/geometric-modeling/notes/bezier-curves/index}
% \input{volume-vi/geometric-modeling/notes/notes-parametric-curves}
% \input{volume-vi/geometric-modeling/notes/notes-bezier}
% \input{volume-vi/geometric-modeling/notes/notes-bsplines}
% \input{volume-vi/geometric-modeling/notes/notes-nurbs}
% \input{volume-vi/geometric-modeling/notes/notes-surface-patches}
% \input{volume-vi/geometric-modeling/notes/notes-continuity}
% \input{volume-vi/geometric-modeling/notes/notes-subdivision}
% \input{volume-vi/geometric-modeling/notes/notes-curvature-analysis}
