
% =========================================================
% Driver Text: Gallian, Contemporary Abstract Algebra (7th ed.)
% File: abstract-algebra/abstract-algebra.tex
% =========================================================

\section{Abstract Algebra}

% =========================================================
% =========================================================

% =========================================================
% I. Integers and Equivalence
% =========================================================

\begin{enumerate}

\item \textbf{Foundations: Integers and Equivalence Relations}

\begin{enumerate}
    \item Modular arithmetic
    \item Equivalence relations and partitions
    \item Functions and mappings
\end{enumerate}

\vspace{0.5em}
\textbf{Structural Theme:}
Congruence and equivalence encode algebraic structure via partitions.
This stage formalizes symmetry at the level of arithmetic.

% =========================================================
% II. Group Theory
% =========================================================

\item \textbf{Groups}

\begin{enumerate}
    \item Definition and basic properties
    \item Subgroups
    \item Cyclic groups
    \item Permutation groups
    \item Isomorphisms
    \item Cosets and Lagrange's Theorem
    \item Normal subgroups and factor groups
    \item Homomorphisms
    \item Fundamental Theorem of Finite Abelian Groups
\end{enumerate}

\vspace{0.5em}
\textbf{Structural Theme:}
Groups formalize symmetry.
Cosets measure deviation from subgroup structure.
Normality permits quotient construction.
Homomorphisms reveal structural preservation.
Classification emerges in the finite abelian case.

% =========================================================
% III. Ring Theory
% =========================================================

\item \textbf{Rings}

\begin{enumerate}
    \item Definition and examples
    \item Integral domains
    \item Ideals and factor rings
    \item Ring homomorphisms
    \item Polynomial rings
    \item Unique factorization domains
    \item Euclidean domains
\end{enumerate}

\vspace{0.5em}
\textbf{Structural Theme:}
Rings generalize arithmetic.
Ideals control structure via quotients.
Factorization encodes algebraic rigidity.
Euclidean domains permit algorithmic structure.

% =========================================================
% IV. Field Theory
% =========================================================

\item \textbf{Fields and Extensions}

\begin{enumerate}
    \item Vector spaces
    \item Field extensions
    \item Algebraic extensions
    \item Finite fields
    \item Geometric constructions
\end{enumerate}

\vspace{0.5em}
\textbf{Structural Theme:}
Fields unify arithmetic and linear algebra.
Extensions enlarge solvability.
Finite fields exhibit deep classification.
Constructibility links algebra and geometry.

% =========================================================
% V. Advanced Topics
% =========================================================

\item \textbf{Special Topics}

\begin{enumerate}
    \item Sylow Theorems
    \item Finite simple groups
    \item Generators and relations
    \item Symmetry groups
    \item Group actions and counting (Burnside)
    \item Coding theory
    \item Galois theory
    \item Cyclotomic extensions
\end{enumerate}

\vspace{0.5em}
\textbf{Structural Theme:}
Group actions connect algebra to combinatorics.
Sylow theory controls finite structure.
Galois theory connects symmetry and solvability.
Cyclotomic extensions bridge algebra and number theory.

\end{enumerate}

% =========================================================
% Structural Remarks
% =========================================================

\vspace{1em}

\begin{remark}
Abstract algebra studies algebraic structures defined by operations
subject to axioms. Each structure (group, ring, field) is a
package consisting of:
\[
\text{Set} + \text{Operations} + \text{Axioms}.
\]
\end{remark}

\begin{remark}
The central construction principle is:
\[
\textbf{Substructure} \longrightarrow \textbf{Quotient Structure}.
\]
Normal subgroups and ideals make quotient constructions possible.
\end{remark}

\begin{remark}[Structural Position]
Abstract algebra in this project builds on:

\begin{itemize}
    \item Set theory and logic (Phase 0)
    \item Linear algebra (vector space structure)
    \item Real analysis (structural proof discipline)
\end{itemize}

Group theory forms the conceptual backbone.
Ring theory extends arithmetic.
Field theory culminates in symmetry of roots (Galois theory).
\end{remark}

% =========================================================
% Input Files (To Be Developed)
% =========================================================

% -----------------------
% Foundations
% -----------------------
\input{volume-vi/book-algebra/abstract-algebra/notes/integers/index}
\input{volume-vi/book-algebra/abstract-algebra/notes/modular-arithmetic/index}
\input{volume-vi/book-algebra/abstract-algebra/notes/induction/index}

% -----------------------
% Groups
% -----------------------
% \input{volume-vi/book-algebra/abstract-algebra/notes/notes-groups}
% \input{volume-vi/book-algebra/abstract-algebra/notes/notes-subgroups}
% \input{volume-vi/book-algebra/abstract-algebra/notes/notes-cyclic-groups}
% \input{volume-vi/book-algebra/abstract-algebra/notes/notes-permutation-groups}
% \input{volume-vi/book-algebra/abstract-algebra/notes/notes-isomorphisms}
% \input{volume-vi/book-algebra/abstract-algebra/notes/notes-cosets}
% \input{volume-vi/book-algebra/abstract-algebra/notes/notes-normal-subgroups}
% \input{volume-vi/book-algebra/abstract-algebra/notes/notes-homomorphisms}
% \input{volume-vi/book-algebra/abstract-algebra/notes/notes-finite-abelian-groups}

% -----------------------
% Rings
% -----------------------
% \input{volume-vi/book-algebra/abstract-algebra/notes/notes-rings}
% \input{volume-vi/book-algebra/abstract-algebra/notes/notes-ideals}
% \input{volume-vi/book-algebra/abstract-algebra/notes/notes-polynomial-rings}
% \input{volume-vi/book-algebra/abstract-algebra/notes/notes-ufd-euclidean}

% -----------------------
% Fields
% -----------------------
% \input{volume-vi/book-algebra/abstract-algebra/notes/notes-field-extensions}
% \input{volume-vi/book-algebra/abstract-algebra/notes/notes-finite-fields}
% \input{volume-vi/book-algebra/abstract-algebra/notes/notes-galois-theory}

% -----------------------
% Special Topics
% -----------------------
% \input{volume-vi/book-algebra/abstract-algebra/notes/notes-sylow}
% \input{volume-vi/book-algebra/abstract-algebra/notes/notes-group-actions}
% \input{volume-vi/book-algebra/abstract-algebra/notes/notes-coding-theory}
